\linespread{1.25}
\chapter{General Overview}


Since the begining of time life has made his way through multiple adverse
conditions. Organisms have evolved from single cells interacting directly with
their environment, to specialised multi cellular entities that respond
specifically to different stimuli. Even though cells can specialise their
functions by differentiating into diverse cell-types, by aggregating into tissues,
generating organs and/or forming a variety of organisms, they all respond to the
same paradigm. DNA is used to save and protect the hereditary information, which
is later transcribed into functional orders in the form of RNA, being most of
the time translated into protein products that execute those final orders.

This process was broadly studied on the twentieth century in order to explain
how life works. DNA was discovered to be the main source of information, in a
double helix format, carrying and saving all necessary instructions to secure
life over generations (\cite{watson1953molecular}). It was discovered a few
years later that the incorporation of the deoxyribonucleotides, from the
triphosphates of deoxyadenosine, deoxyguanosine, deoxycytidine and thymidine,
into the deoxyribonucleic acid was catalyzed by an enzyme, the DNA polymerase,
which was purified at that time directly from \textit{Escherichia coli}
cell-free extracts (\cite{lehman1958enzymatic}). 

Proteins were further studied as the final product of DNA's information
translatation, helped by the discovery of the green fluorescent protein (GFP). A
naturally fluorescent protein present in the \textit{Aequorea Victoria}
jellyfish (\cite{shimomura1962extraction}). This allowed the generation of
proteinic chimeras, detection and real-time tracking of the products. 

Another type of enzyme that would revolutionise Molecular Biology were the
Endonucleases. Discovered in the early 60s by \cite{arber1962host} and
recognised as "molecular scissors" protecting bacteria from bacteriophages.These
restriction enzymes work in conjuction with methylases protecting the DNA of the
bacterium. The first Type IIS enzyme described was HindII isolated for the first
time from \textit{Haemophilus influenzae Rd} \cite{smith1973suggested}, which
also described the 6 bp site-specific cleavage. The understanding of how
restriction enzymes cut DNA, and how host DNA tries to protect itself are the
basis of modern genetic engineering.

Thanks to the advances in DNA cleavage, recombinant DNA molecules were created
for the first time by the fusion of genetic material from more than one species.
Here the DNA from two viruses was cut creating “sticky ends” and the ends
annealed, finishing with the addition of DNA ligase for sealing it.
\cite{jackson1972biochemical}. The aquired knowledge was further expanded by the
corroboration of the previous experiments via gel electrophoresis and testing
them in SV40 viral genome \cite{danna1971specific}, work that was vital for the
generation of genomic mapping.

The introduction of recombinant DNA into living organisms was achieved by the
work \cite{cohen1973construction}, proving that DNA can replicate naturally,
despite being artificially introduced in another organism. Establishing the
principles of modern genetical engineering.

Most previous achievements would not have been posible without the discovery of
\textit{in vitro} translation via the usage of cell-extracts.
\cite{miller1970visualization} demontstrated that RNA and protein synthesis can
occur almost simultaneously in the cell, and that this process can be completly
replicated in a test tube. Even though the coupling of transcription and
translation accellerate the product production, proteins still need to acquire
their secondary, tertiary or even quaternary conformation in order to become
functional. Even sometimes needing post-translational modifications to work.  

As it can be observed, the usage of cell-free technology is not a new technic
neither Synthetic Biology. For instance, one of the first mentions of Synthetic
Biology as a term can be tracked way back to \cite{leduc1912biologie}, who sought to
synthesise life "by directing the physical forces which are its cause".


Cell-free expression systems are a powerful tool in applications where cell
based expression systems are too variable, slow, difficult to store or
prohibitive due to legislation on the release of genetically modified organisms.
Cell-free systems (CFS) are not without their limitations; since their very
nature is to be a defined, precisely controlled system that lacks
self-replicating functionality, they depend on a limited quantity of supplies,
such as ATP, amino acids and nucleotides \cite{kwon2015high}. Consequently,
in-vitro synthesis of protein is limited by two dimensions: the overall maximum
protein amount that can be synthesized with the supplied components, and the
time during which the CFS stays active before background processes and chemical
deterioration have used up supplies or inactivated the system
\cite{carlson2012cell,bernhard2013cell,kwon2015high} One of the most promising
applications of CFS is the possibility to design cell-free biosensors, which do
not present a living prospect subject to the current GMO legal regulations and
offer an alternative to standard analytical techniques, such as ICP-MS and
ICP-OES. However, cell-free biosensors that rely on protein expression also
present the common limitations of CFS, such as a long response time and a lack
of precursor regeneration. Sensors that could function without the need for
protein expression would drastically reduce the requirements for such a system.
Paige et al., (2011) designed synthetic fluorescent RNA aptamers that mimic
fluorescent proteins without the need for translation. The secondary structure
of these aptamers presents a loop in which a fluorophore is trapped, conferring
the same intramolecular immobilization that confers fluorescence in fluorescent
proteins (Paige et al. 2011; Strack et al. 2013). These RNA aptamers also suffer
from several limitations as incorrect folding is observed under non optimal
thermal or ionic conditions (Autour et al. 2016). Improved aptamers with a
stronger secondary structure, such as Broccoli and iSpinach, show increased
stability under non-optimal conditions (Filonov et al. 2014, Autour et al.
2016). Both aptamers are based on the 95 base core sequence of Spinach2 and can
be flanked by a tRNA scaffold, increasing the stability of their secondary
structures (Filonov et al. 2014, Autour et al. 2016, Strack et al. 2013).
However, these were designed for in vivo use (e.g. analysing RNA quantities) and
have never been tested as outputs for a cell-free sensor. Such a sensor could
require only transcription, not translation, hence could be much faster and
simpler. On the other hand, the simplicity of CFS can become a disadvantage when
complex processes need to beperformed. For instance, the high dependence on E.
coli as a model organism generates a problem, since genes or genetic pathways
from non-model organisms or obtained by metagenomic studies, may fail to be
expressed in standard organisms. Moore et al., (2016 and 2017) have attempted to
solve this issue by using non-model bacteria, such as Bacillus or Streptomyces,
for the generation of new cell-free TX/TL systems. This approach can be applied
to other organisms, such as extremophiles, in order to use their unique
capacities. For example, Cupriavidus metallidurans possesses the capability to
tolerate, remove and degrade diverse environmental pollutants(Millacura et al.,
2017). A cell-free system based on Cupriavidus metallidurans extract might show
increased tolerance for heavy metals and might be superior for heavy metal
sensors. Another challenge for CFS is to analyze multiple variables
simultaneously. As they rely on the normal cell processing/synthesis machinery
(interaction between transcription factors, polymerases, ribosomes, and other
diverse macromolecules), they suffer the same issues as a living organism.
Furthermore, when complex interactions are carried out, problems may arise due
to kinetic mismatches, lack of oscillation or Boolean/Reversible designs (Guz et
al., 2016). The generation of cell-free systems that respond to variables using
totally synthetic processing machinery seems to be one of the most promising
approaches. There have been some attempts to generate logic gates without
mimicking normal cell machinery (Bordoy et al., 2016, Chatterje et al., 2016,
Kim et al., 2014 and Zhang et al., 2016), however, due to their dependence on
translation and/or the need for manually addition of foreign oligonucleotides
(Kim et al., 2014) they still show slow response. Additionally, some of them
rely on recombination processes that make their implementation even more
complicated than under in vivo conditions (Zhang et al., 2016). In electronics,
a single circuit accepts one or more binary inputs to generate one or more
binary outputs. There have been many attempts to replicate such circuits using
in vivo genetic networks. A typical biological logic gate consists of an output
macromolecule that is produced only if the corresponding pattern of inputs is
present, inputs that are commonly associated with the presence of
transcriptional regulators, transcription factors, polymerases or other
macromolecules \cite{silva2008mining} . Here we propose an alternative approach,
BioLogic, which decomposes a large circuit into a collection of small
subcircuits, solved in parallel. Rather than having a single type of cell (or
genetic material) doing the computation, we have separate versions (subcircuits)
each reacting to a different combination of inputs and generating the desired
response by combination of each final output. For instance, if each input bit is
considered in two forms, ZERO and ONE, each of which is essential to certain
output agents, any arbitrary pattern of outputsfor any pattern of inputs could
be generated, making all kinds of binary operation possible A further proposal
is to achieve this goal by using just a transcriptional genetic network, without
need for translation. Each agent in this case, rather than a cell, is a plasmid
or double stranded DNA that does not possess functional transcriptional networks
(Figure 1). As cell-free systems can assemble DNA parts in vitro at very low
cost by using cell extracts instead of enzyme cocktails (Casini et al., 2015),
and Golden Gate Assembly is an excellent tool to assemble multiple DNA fragments
in a defined linear order, by using type IIS restriction enzymes (Engler et al.,
2008), these techniques may be combined to generate a new and inexpensive in
vitro system, which through DNA assembly and processing (hereafter DNALogic)
would be able to sidestep common problems associated with the use of live cell
based systems. As a "triggering" step is vital for the DNALogic processing
machinery, the novel CRISPR/Cpf1 technology (or similar) will be used to
generate the respective recognition and cutting steps. CRISPR/Cpf1 generates 5'
overhanging cuts, producing 8 nt sticky ends and cleaving at the 14th base from
the PAM site TTN (Lei et al., 2017 and Li et al., 2016). Due to its similarities
to Type IIS restriction enzymes, Cpf1 can also be used to generate a new kind of
Golden Gate Assembly, which would not rely on the presence or absence of cutting
sites, but on gRNA. The gRNA expression system can be combined with the
recognition step, working as a transduction system between the detection and
processing steps (Figure 1a). This CRISPR triggered sequence recognition,
followed by cutting and ligation steps, would result not only in a fully
functional transcriptional unit, but also in a different double stranded genetic
sequence, accomplishing the parameters for synthetic memory generation explained
and demonstrated in vivo by Siuti et al. (2013). This creates DNA memory that
can not only be later amplified and processed by sequencing and/or PCR (later
response), but that can also generate RNA expression of a fluorescent aptamer or
a further gRNA if there is formation of a proper transcription unit (immediate
response).Figure 1. Proposed DNALogic system. A) crRNA expression is regulated
by inducible promoters B) Ribonucleoprotein formation is followed by recognition
and cutting of the target DNA C) Ligation of the overhangs produced by the
Ribonucleoprotein generates a transcriptional unit D) Expression of the
transcriptional unit generates a functional fluorescent RNA aptamer.Aims: This
work has as its general aim the generation of a modular cell-free system capable
of not only sensing variables present in the surrounding environment, such as
heavy metals, but that is also able of analyzing them by using a totally
synthetic in vitro logic gate. Specific aims are: 1. Testing fluorescent RNA
aptamers as output signals for cell-free sensors. 2. Developing Cpf1-based
signal processing technology for use in cell-free systems. 3. Creation of an
adaptable DNALogic gate, which allows memory and more complex analysis in cell
free systems. 4. Testing Cupriavidus metallidurans extracts as basis for
metal-sensing cell free systems using endogenous and exogenous metal sensing
promoters. 5. Integrating these to develop a modular cell-free sensing system
with versatile in-vitro signal processing.
